### 2021-03-28 Manuscript.Rmd
Finalization

---
title: "Storm Data Analysis - Week 4"
author: "Marco Baciarello"
date: "3/17/2021"
output: html_document
  fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)

suppressMessages(library(tidyverse))
suppressMessages(library(dtplyr))
suppressMessages(library(data.table))
# library(skimr)
```

```{r function-definitions}

# WARNING: operates by reference, modifies argument data.table Adds columns for
# ranks based on event count, damage and injuries/fatalities. Useful to quickly
# re-rank tables with different elements or filters based on year.
rank_DT <- function(DT) {
    DT[, .(
            counter = .N, # Count elements in group
            # Total monetary damages per group
            total_dmg = sum(tot_dmg, na.rm = TRUE),
            # Total fatalities
            fatalities = sum(FATALITIES, na.rm = TRUE),
            # Injuries
            injuries = sum(INJURIES, na.rm = TRUE)
            ),
        # Grouping by event type
        by = event][
            ,
            # Compute a "percent rank": rank events by number of
            # occurrences (desc), then assign a percentile by
            # dividing rank by denominator (i.e. nrows)
            # Note double dot syntax to access calling scope
            `:=`(perc_rank = frankv(counter)/..denominator,
                 # Same for monetary damage
                 dmg_rank = frankv(total_dmg)/..denominator,
                 # And life damage
                 inj_rank = frankv(injuries)/..denominator,
                 fat_rank = frankv(fatalities)/..denominator
                 )
            ][
                # Filter for frequency rank >=90th
                perc_rank >= 0.9
            ][
                order(-counter)
            ]
}

# Min-max normalization
normalizer <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
  }

```

## Abstract

Blah blah

## Introduction

We analyze this, that and something else...

## Data Processing

First, we download the data file and load it into an R object. We're using a `tidyr`-centric approach, leveraging on the full suite of library found in package `tidyverse`.

```{r data-prep, message=FALSE, warning=FALSE, cache = TRUE}

# Check if data dir exists -> create one
if (!file.exists("data")) {
    dir.create("./data")
}

# Get the data
if (!file.exists("data/stormdata.csv.bz2")) {
    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", "data/stormdata.csv.bz2")
}

# Read data in. fread seems way faster than other solutions.
rdata <- fread("data/stormdata.csv.bz2", data.table = FALSE)
# Convert to "lazy data.table" which will accept dplyr functions
data <- lazy_dt(rdata)

```

### Time formatting

We create 2 new table columns; one is a properly-formatted `POSIXct` column including date and time information from columns `BGN_DATE` and `BGN_TIME`. We will ignore end times of phenomena for now.

```{r column-formatting}

# TECH NOTE: We're using the `dtplyr` interface to use `tidyverse` grammar but
# run data.table code, for increased performance on large tables.

# Format data
fdata <- data %>%
    mutate(
        # Paste date and time For some reason, data includes a fixed time 00:00
        # which we don't need, so we split and get the first fragment
        temp_datetime = paste(strsplit(BGN_DATE, " ")[[1]][1], BGN_TIME),
        # Convert to datetime; 
        utc_datetime = as.POSIXct(
            temp_datetime,
            format = "%m/%d/%Y %H%M"),
        # correct timezones (there are older/unsupported abbreviations)
        c_timezone = recode(TIME_ZONE,
                            CDT = "US/Central",
                            CST = "US/Central",
                            UNK = "UTC",
                            MDT = "US/Mountain",
                            AST = "US/Eastern",
                            ADT = "US/Eastern",
                            CSt = "US/Central",
                            CSC = "UTC",
                            SCT = "US/Central",
                            ESY = "US/Eastern",
                            SST = "US/Samoa",
                            AKS = "US/Alaska",
                            GST = "US/Central"
                            ),
        # # Convert each time to its own time zone
        datetime = lubridate::force_tzs(utc_datetime, tzones = c_timezone),
        # Force POSIXct format, otherwise data won't be usable in data.table
        # package functions
        begin_time = as.POSIXct(datetime),
        # Finally, change EVTYPE into a factor ordered by frequency (may be
        # CPU-intensive)
        event = fct_infreq(as.factor(str_to_lower(EVTYPE)))
    ) %>%
    # Need to drop non-POSIXlt columns for data.table compatibility
    select(-contains("datetime")) %>%
    # End dtplyr "session," translate into data.table code and run
    as_tibble()

```

### Property and agricultural damages

A second column will be created in order to reconstruct information on property damage: columns `PROPDMGEXP` specifies an "exponent" for `PROPDMG`, in the form of a mathematical suffix: K for thousands, M for millions, and so on... Values can also be integers, in which case they seem to indicate that the reported damage value is to be multiplied by 10 (see full legend [here](ane.st/HandleExponent6c0f3 "Legend for EXP values"). We will compute a column containing the actual amount as an integer or double. Same operation will be performed for `CROPDMG` and `CROPDMGEXP`.

```{r exponents}
unique(fdata$PROPDMGEXP)
unique(fdata$CROPDMGEXP)
```

We will transform them into the actual multiplier so we can compute actual damages. It is known that there are a few malformed values [*e.g.*: empty strings ("")]; they will be set to 0. According to the above analysis, malformed values are negligible both in numbers and absolute values.

```{r exp-to-multipliers}

# Cannot use dtplyr here because of error: Can't rename variables in this context.

temp_data <- fdata %>%
    mutate(
        year = as.integer(lubridate::year(begin_time)),
        prop_mult = case_when(
            PROPDMGEXP %in% c(0, 1, 2, 3, 4, 5, 6, 7, 8) ~ 10,
            PROPDMGEXP %in% c('m', 'M') ~ 1000000,
            PROPDMGEXP %in% c('k', 'K') ~ 1000,
            PROPDMGEXP %in% c('b', 'B') ~ 1000000000,
            PROPDMGEXP %in% c('h', 'H') ~ 100,
            PROPDMGEXP == '+' ~ 1,
            # Vector must be double because of limitations in `case_when`, so we
            # can't set a value to NA; use dummy code and convert later.
            TRUE ~ -999999
        ),
        crop_mult = case_when(
            PROPDMGEXP %in% c(0, 1, 2, 3, 4, 5, 6, 7, 8) ~ 10,
            CROPDMGEXP %in% c('m', 'M') ~ 1000000,
            CROPDMGEXP %in% c('k', 'K') ~ 1000,
            CROPDMGEXP %in% c('b', 'B') ~ 1000000000,
            CROPDMGEXP %in% c('h', 'H') ~ 100,
            CROPDMGEXP == '+' ~ 1,
            # Vector must be double because of limitations in `case_when`, so we
            # can't set a value to NA; use dummy code and convert later.
            TRUE ~ -999999
        ),
        # Change dummy value into NA
        across(ends_with("_mult"), ~na_if(., -999999)),
        crop_dmg = CROPDMG * crop_mult,
        prop_dmg = PROPDMG * prop_mult,
        tot_dmg = prop_dmg + crop_dmg
    )

# Switch back to dtplyr syntax and clean up
temp_data <- lazy_dt(temp_data)

# Filter out events which did not cause any injuries, fatalities or monetary
# damages
mdata <- temp_data %>% filter(
  (INJURIES > 0) | 
    (FATALITIES > 0) |
    (tot_dmg > 0)
) %>%
  # Apply data.table code
  as_tibble()

```

```{r invalid-exps, eval=FALSE, include=FALSE}

# DEBUG: this chunk not to be run nor shown when knitting.

# Property damage
inv_prop_exp <- filter(temp_data, PROPDMG != 0, prop_mult == -999999) %>%
  select(PROPDMG, PROPDMGEXP, prop_mult, BGN_DATE) %>%
  as.data.frame() %>%
  nrow()

inv_crop_exp <- filter(temp_data, CROPDMG != 0, crop_mult == -999999) %>%
  select(CROPDMG, CROPDMGEXP, crop_mult, BGN_DATE) %>%
  as.data.frame() %>%
  nrow()

# Clean up
# rm(temp_data)
```

### Injuries and fatalities

These data do not seem to be needing any significant adjustments:

```{r inj-fat-describe}

skimr::skim(mdata$INJURIES)
cat("\n\n\n")
skimr::skim(mdata$FATALITIES)

```

### Selection of variables of interest

According to the requirements for this report, we will be describing associations between weather events and human and economical damages. After preparatory operations performed above, we will now drop variables which are not of interest to these analyses, in order to make data more easily readable and reduce resource requirements.

```{r data-column-thinning}

# Select useful columns
lean_data <- mdata %>%
    select(REFNUM, event, year, c_timezone,
           FATALITIES, INJURIES, tot_dmg)

```

### Selection of events

The vast majority of events have caused no damage to people or property. More than 90% of event *types* are not associated with damage, and many of them have occurred just once since 1950. This is mostly due to numerous, uncorrected mistakes made in data entry over the first ⅔ of the database's life.

```{r evtype-selection}

# Make a copy of data and cast as DT
sel_data <- as.data.table(lean_data)
# Apply final transformations to DT
sel_data2 <- sel_data[, .(
    counter = .N, # Count elements in group
    # Total monetary damages per group
    total_dmg = sum(tot_dmg, na.rm = TRUE),
    # Total fatalities
    fatalities = sum(FATALITIES, na.rm = TRUE),
    # Injuries
    injuries = sum(INJURIES, na.rm = TRUE)
  ),
  # Grouping by event type
  by = .(event)][
    ,
    # Compute a "percent rank": rank events by number of
    # occurrences (desc), then assign a percentile by
    # dividing rank by denominator (i.e. nrows)
    # Note double dot syntax to access calling scope
    `:=`(perc_rank = frankv(counter)/.N,
         # Same for monetary damage
         dmg_rank = frankv(total_dmg)/.N,
         # And life damage
         inj_rank = frankv(injuries)/.N,
         fat_rank = frankv(fatalities)/.N
    )
  ][
    # Filter for frequency rank >=95th
    perc_rank >= 0.80
  ][
    order(-counter)
  ]

```

```{r ev-selection-comparison}

gdata <- as.data.table(temp_data)
# With DT syntax, group and compute relevant sums
totals <- gdata[,.(sum_inj = sum(INJURIES, na.rm = T),
               sum_fat = sum(FATALITIES, na.rm = T),
               sum_dmg = sum(tot_dmg, na.rm = T))]

selected <- sel_data2[,.(sum_inj = sum(injuries, na.rm = T),
               sum_fat = sum(fatalities, na.rm = T),
               sum_dmg = sum(total_dmg, na.rm = T))]

# Proportion of damages in selected events as opposed to total
round((selected/totals), digits = 2)
```

Events below the 80^th^ percentile of frequency are mostly irrelevant to the analysis. To confirm this, we calculated total damages, injuries and fatalities from the global data and from the selected data. The ratios of the selected data to the global data is near 1 (see above), despite the selected data table having only a fraction of the event types (`r length(unique(gdata$event))` *vs.* `r length(unique(sel_data2$event))`).

### Aggregation of events by type

We then aggregate event types into arbitrary categories, both for clarity and because there is artificial variability in the form of spelling errors and variations (*e.g.*: `hail`, `marine hail`, `hail/wind`, `small hail` and so forth...)

Nested loops is our way of going through a factor variable and associating each value to a value in a new factor variable (many-to-one relationship.). The need for loops seems to be related to the fact that these vars are stored as factors, and the lack of a specific function in the forcats package; fct_collapse() might have been one solution, but it takes the exact factor level names—since there were almost 1000 different possible levels, it would have been impractical to call this function.

```{r event-aggregation, warning=FALSE}

# Functions to aggregate (collapse) factor levels require specific values, specified in a named vector. Here, we use "data mining" to generate such vector.

# Create regexp patterns to match to. The following is also the priority order
# with which events will be assigned to a category.
new_fcts <- list(
  hail = c("hail"),
  tornado = c("torn", "nado"),
  tsunami = c("tsun"),
  stormy = c("storm*", "thunder*", "tstm", "rain", "precip", "hurric", "spout"),
  lightning = c("lightn"),
  flood = c("flood", "seiche", "fld"),
  cold = c("snow", "cold", "ice", "sleet", "^*inter", "blizz", "avalan"),
  heat = c("heat", "hot", "summ", "warm"),
  windy = c("wind", "devil", "b*rst"),
  fog = c("fog"),
  maritime = c("sea*", "ocean*", "mari[nt]*",
               "curr*", "tid*", "surf*", "wav", "swell"),
  drought = c("drought"),
  fire = c("fire", "blaze", "smok"),
  landslide = c("slide")
)

# Initialize list, vector and counter for the loop
old_events <- vector("list", 14)
new_events <- c()
i <- 1
```

We create a list of possible synonims for weather events. In the inner loop, the factor variable is run (using `stringr::subset()` sequentially against each list element and its children. Where there's a match, the function returns the full value of the variable. We aggregate these values in a vector. In the outer loop, the unique values (i.e., factor levels) are appended to a second list. The second list has a list-element for each of the new factor levels; the old factor levels are thus stored as children of the new ones, based on matching with our criteria.

Once complete, the list _could_ be used as a named vector. However, for some reason we had to insert each element manually using list$element syntax.

```{r factor-loops}
suppressWarnings(
    # For every new event category
    for (i in seq(1:length(new_fcts))) {
        # Prepare storage array
        old_instances <- c()
        k <- 1
        # Loop over elements in list element
        for (k in seq(1:length(new_fcts[[i]]))) {
            # Match old event instances
            old_instances <- c(old_instances,
                               str_subset(sel_data2$event, new_fcts[[i]][k])
                                )
            k <- k + 1
        }
        # Add to list
        old_events[[i]] <- unique(old_instances)
        i <- i + 1
    }
)
# Name (tag) list elements
names(old_events) <- names(new_fcts)

# TODO: switch to using list names instead of manual entry of new factor levels
sel_data2$event2 <- fct_collapse(
  sel_data2$event,
  hail = old_events$hail,
  tornado = old_events$tornado,
  tsunami = old_events$tsunami,
  stormy = old_events$stormy,
  lightning = old_events$lightning,
  flood = old_events$flood,
  cold = old_events$cold,
  heat = old_events$heat,
  windy = old_events$windy,
  maritime = old_events$maritime,
  fog = old_events$fog,
  drought = old_events$drought,
  fire = old_events$fire,
  landslide = old_events$landslide,
  other_level = "Other events"
) %>%
  # Adjust spelling/capitalization of factor levels
  fct_recode(
    "Hail" = "hail",
    "Winds" = "windy",
    "Tornadoes" = "tornado",
    "Floods" = "flood",
    "Lightnings" = "lightning",
    "Cold weather" = "cold",
    "Storms/Rains" = "stormy",
    "Oceanic events" = "maritime",
    "Fires" = "fire",
    "Droughts" = "drought",
    "Hot weather" = "heat",
    "Fog" = "fog",
    "Landslides" = "landslide",
    "Tsunamis" = "tsunami"
  ) %>%
  # Reorder according to frequency, desc
  fct_infreq()

```

The choice of the fourteen categories was arbitrary, based on observation of event types appearing in the selected data. There are limitations to this approach, namely: events belonging to more than one category; and possible oversight of extremely rare events.

```{r memory-release-cleanup}

# Clean up large dataframes, no longer used
erase_us <- c("data", "fdata", "gdata", "mdata", "rdata")
rm(list = erase_us)
```

```{r figure-captions}
fig.1.caption = "**Fig. 1:** Figure 1"
```

## Results

Data on agricultural (crop) and property damages show unequivocal predominance of flooding phenomena as causes. At `r scales::dollar(dmgdata$total_dmg[1])`, overflowing of oceans, lakes and rivers caused about 4 times as much damage as storms and other rain phenomena, a distant second in ranking. One can easily imagine damages from disasters such as hurricanes Katrina and Sandy are the primary drivers of this result.

```{r damage-plot, fig.cap=fig.1.caption}

# Make a copy of data to be sorted by damage total, descending, for use in
# damage-related plots.
dmgdata <- sel_data2 %>%
    group_by(event2) %>%
    summarise(total_dmg = sum(total_dmg),
              total_inj = sum(injuries),
              total_fat = sum(fatalities)
              ) %>%
    mutate(event2 = as.character(event2)) %>%
    arrange(desc(total_dmg)) %>%
    as_tibble()

### Damage plot

dmgdata %>%
  ggplot(aes(
    x = reorder(event2, -total_dmg),
    y = total_dmg)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(prefix = NULL,
                                  scale = 1e-9,
                                  largest_with_cents = 1),
    name = "Total estimated damages caused\n(in billions of dollars)\n"
  ) + 
  scale_x_discrete("\nWeather event category (1950–2013)") +
  coord_cartesian(ylim = c(0, 5e10)) +
  annotate("segment", x= 2.2, xend = 1.5, y = 50e9, yend = 52e9,
           arrow = arrow(angle = 20,
                         length = unit(0.15, "inches"),
                         type = "closed")
  ) + 
  annotate("text",
           x = 4.3, y = 49e9,
           label = "Plot clipped for readability.\nActual value: $147 billion") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
#####

```

```{r injuries-plot}
### Injuries plot

dmgdata %>%
    ggplot(aes(
        x = reorder(event2, -total_inj),
        y = total_inj)) +
    geom_col() +
    scale_x_discrete("\nWeather event category (1950–2013)") +
    scale_y_continuous(name = "Related injuries reported\n",
                       breaks = seq(0, 1e5, 1e4),
                       labels = scales::label_comma()) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 70, hjust = 1))

```

```{r fatalities-plot}
### Fatalities plot

dmgdata %>%
    ggplot(aes(
        x = reorder(event2, -total_fat),
        y = total_fat)) +
    geom_col() +
    scale_x_discrete("\nWeather event category (1950–2013)") +
    scale_y_continuous(name = "Related fatalities reported\n",
                       #breaks = scales::extended_breaks(),
                       breaks = seq(0, 5e3, 1e3),
                       labels = scales::label_comma()) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 70, hjust = 1))
```

## Conclusions